### 2021-11-12 rmd/summary_slides.Rmd
add: rmd files

---
title: "8/6 Meeting"
author: Akira Terui
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

## Captive breeding

- Form of conservation of critically endangered species
- Restore declining natural populations
- Effective in certain circumstances

![](image/img-florida-panther.jpg)
Photo - Wikipedia

- Increased awareness of "downside"

## Knowledge gap

Community-level impact of captive breeding is unknown

- Effects spread through species interactions?
- Enhancement with captive-reared organisms sometimes excessive
- Fish stocking is an example

## Question

How does fish stocking influence long-term stability of fish communities?

- Theory: community model with a Ricker formula + stocking
- Data: empirical test with long term data

## Question

How does fish stocking influence long-term stability of fish communities?

- **Theory: community model with a Ricker formula + stocking**
- Data: empirical test with long term data

## Theory

Community Ricker model

$$
N_{i,t+1} = N_{i,t}exp(r_i[1-\frac{\sum_{j=1}^n\alpha_{ij}N_{j,t}}{K}])
$$
$N$ - density  
$r$ - intrinsic population growth rate  
$K$ - carrying capacity  
$\alpha$ - competition coefficient  
$i,j$ - species  
$t$ - time

## Theory

Community Ricker model + stocking

$$
N_{i,t+1} = (N_{i,t} + s_t)exp(r_i[1-\frac{\sum_{j=1}^n\alpha_{ij}N_{j,t} + \alpha_{i1}s_t}{K}])
$$

$N$ - density  
$r$ - intrinsic population growth rate  
$K$ - carrying capacity  
$\alpha$ - competition coefficient  
$i,j$ - species  
$t$ - time

## Theoretical prediction

Community-level CV increased with stocking even with weak species interactions

```{r figure1, fig.width = 7, fig.height = 4}
source("code_theory/figure_theory.R")
```

## Question

How does fish stocking influence long-term stability of fish communities?

- Theory: community model with a Ricker formula + stocking
- **Data: empirical test with long term data**

## Data

- Fish data collected in protected watersheds in Hokkaido
- Data compiled for 1999-2019

---

```{r figure2, fig.width = 8, fig.height = 6}
source("code_empirical/figure_time_series.R")
```

## Results

Community-level CV increased with stocking

```{r figure3, fig.width = 7, fig.height = 5}
source("code_empirical/figure_cv_stock.R")
```