### 2022-03-03 rmd/title.tex
edit: suppl info

%\begin{titlepage}
\begin{center}
\LARGE{\textbf{Supplementary Materials for:}}\\[7.5mm]
\LARGE{Intentional stocking undermines ecological stability}\\[7.5mm]
\large{Akira Terui, Hirokazu Urabe, Masayuki Senzaki, Bungo Nishizawa}\\[7.5mm]
\large{Correspondence to: hanabi0111@gmail.com}\\[15mm]
\end{center}
\begin{flushleft}
\textbf{This PDF file includes:}
\begin{itemize}
\item Supplementary text
\item Tables S1-10
\item Figures S1-S11
\item References
\end{itemize}
\end{flushleft}
% \end{titlepage}

\newpage