### 2022-03-12 rmd/title.tex
Squashed commit of the following:
commit 42b7838e55c1587f6a744b0a978c367eac453d43
Author: AkiraTerui <hanabi0111@gmail.com>
Date:   Sat Mar 12 11:50:40 2022 -0500

    edit: affiliation

commit 1933aaca4b7cc7f4d29771b59b45cf512e403b48
Author: aterui <hanabi0111@gmail.com>
Date:   Thu Mar 10 22:06:24 2022 -0500

    edit:  manuscript

commit 0884ad16cc1eb71df364fe1aaf0e7cec0dadd794
Author: aterui <hanabi0111@gmail.com>
Date:   Thu Mar 10 21:32:32 2022 -0500

    edit: manuscript

commit db4e8b6bab9c85a4ba5ba4767ac5442fcf493553
Author: AkiraTerui <hanabi0111@gmail.com>
Date:   Thu Mar 10 19:04:2

%\begin{titlepage}
\begin{center}
\LARGE{\textbf{Supplementary Materials for:}}\\[7.5mm]
\LARGE{Intentional release of native species undermines ecological stability}\\[7.5mm]
\end{center}
\begin{flushleft}
\textbf{This PDF file includes:}
\begin{itemize}
\item Methods
\item Supplementary text
\item Tables S1-10
\item Figures S1-S11
\item References
\end{itemize}
\end{flushleft}
% \end{titlepage}

\newpage