### 2022-07-12 rmd/ppt_test.Rmd
add: deterministic model

---
date: "`r Sys.Date()`"
author: "Your Name"
title: "Untitled"
output: 
  officedown::rpptx_document:
    reference_doc: alison-template.pptx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.cap = TRUE)
library(officedown)
library(ggplot2)
library(rvg)
```

## Table

```{r layout='Two Content', ph=officer::ph_location_left()}
head(cars)
```

```{r layout='Two Content', ph=officer::ph_location_right()}
head(iris)
```

## Slide with Plot

```{r plot1}
dml(plot(pressure))
```

## Slide with ggplot

```{r layout='Title and Content', ph=officer::ph_location_type(type="body")}
mygg <- ggplot(pressure, aes(x = temperature, y = pressure)) +
  geom_point()
dml(ggobj = mygg)
```

## Slide with ggplot

```{r ph=officer::ph_location_fullsize()}
mygg <- ggplot(pressure, aes(x = temperature, y = pressure)) +
  geom_point()
dml(ggobj = mygg)
```

## ggplot in left column

```{r ph=officer::ph_location_left()}
mygg <- ggplot(pressure, aes(x = temperature, y = pressure)) +
  geom_point()
dml(ggobj = mygg)
```

## ggplot in right column

- Some
- Bullets

```{r ph=officer::ph_location_right()}
mygg <- ggplot(pressure, aes(x = temperature, y = pressure)) +
  geom_point()
dml(ggobj = mygg)
```

## ggplot in both columns

```{r ph=officer::ph_location_left()}
dml(ggobj = mygg)
```

```{r ph=officer::ph_location_right()}
dml(ggobj = mygg)
```

## custom plot locations

```{r ph=officer::ph_location(width = 2, height = 2, left = 2, top = 2)}
dml(plot(pressure))
```

```{r ph=officer::ph_location(width = 3, height = 3, left = 4, top = 4)}
dml(plot(pressure))
```

## Non-editable plot

```{r}
dml(plot(pressure), editable = FALSE)
```