### 2022-09-28 rmd/README.Rmd
readme

---
title: README
output: github_document
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile,
                          encoding = encoding, 
                          output_dir = here::here())
      })
---

```{r, include=F}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

## Article Information

Title: Intentional release of native species undermines ecological stability
Authors: Akira Terui, Hirokazu Urabe, Masayuki Senzaki, Bungo Nishizawa

## Contents

```{r}
source(here::here("code/set_readme.R"))
knitr::kable(tbl_readme)
```

## Session Information

```{r}
sessionInfo()
```