### 2022-12-26 rmd/title.tex
edit: minor change

%\begin{titlepage}
\begin{flushleft}
\vspace*{5mm}
\LARGE{\textbf{Supplementary Materials for:}}\\[2.5mm]
\LARGE{Intentional release of native species undermines ecological stability}\\[7.5mm]
\large{Akira Terui, Hirokazu Urabe, Masayuki Senzaki, Bungo Nishizawa}\\[7.5mm]
\large{Correspondence to: Akira Terui hanabi0111@gmail.com}\\[15mm]
\textbf{This PDF file includes:}
\begin{itemize}
\item Supplementary text
\item Tables S1 -- 8
\item Figures S1 -- S17
\item References
\end{itemize}
\end{flushleft}
% \end{titlepage}

\newpage